---
title: "Final exam_2016140072_Jeeyoon_Chang"
output:
  html_document:
    df_print: paged
---

# Final exam_Jeeyoon_Chang

# Q1.
From   the   dataset   `heights`   in   the   `dslabs`   package,   please   describe   the distribution   of   male   and   female   heights.   If   you   pick   a   female   at   random,   what   is the   probability   that   she   is   61   inches   or   shorter?

```{r}
# The distribution of height between male and female.
# The overall (average) height of male is taller than female.
library(dplyr)
library(ggplot2)
library(dslabs)
library(ggplot2)
data(heights)

# violin plot
heights %>% 
  ggplot( aes(sex, height, color = sex)) +
  geom_violin() + geom_boxplot()
```

```{r}
# density plot
heights %>% 
  ggplot( aes(height, color = sex)) +
  geom_density()
```
```{r}
# probability of random female taller than 61 inches
F <- heights %>% filter (sex == "Female")
F61 <- F %>% filter (height <= 61 )
Answer = 29/238

# [1] 0.1218487
```

# Q2.
 For   American   Roulette,   there   are   19   reds,   16   blacks   and   3   greens.   The   payout for   winning   on   green   is   15   dollars.   You   create   a   random   variable   that   is   the   sum   of your   winnings   after   betting   on   green   1000   times.   Start   your   code   by   setting   the seed   to   1.   Describe   your   random   variable   (e.g.   the   expected   value,   the   standard error).   Then,   create   a   Monte   Carlo   simulation   that   generates   1,000   outcomes   of   a random   variable,   and   then   describe   your   result. 
 
```{r}
# The random variable of the sum of your winninng
set.seed(1) 
n <- 1000
X <- sample(c(15,-15), n, replace = TRUE, prob=c(3/38, 35/38))
S <- sum(X)

# S = -12240

# Expected value
B <- 10^6
x <- sample(c(15,-15), B, replace = TRUE, prob=c(3/38, 35/38))
mean(x)

# Standard error

n <- 1000
sqrt(n) * 2 * sqrt(35*3)/38

# Monte Carlo 
```


# Q3.
 From   the   poll   example,   we   will   create   a   Monte   Carlo   simulation   for   p   =   0.45. You   will   compare   the   sampling   size   (N)   for   10,   1000,   and   the   repeat   size   (B)   for 100,   10000.   So   you   should   have   four   combinations   (10   N   x   100   B,   1000   N   x   100   B, 10   N   x   10000   B,   1000   N   x   10000   B).   Please   describe   your   Monte   Carlo   simulation results,   and   compare   four   combinations.
 
```{r}
B <- 10000
N <- 1000
p <- 0.45
x_hat_1 <- replicate(B, {
  x <- sample(c(0,1), size = N, replace = TRUE, prob = c(1-p, p))
  mean(x)
})
mean(x_hat_1)
sd(x_hat_1)

```

```{r}
B <- 10000
N <- 10
p <- 0.45
x_hat_2 <- replicate(B, {
  x <- sample(c(0,1), size = N, replace = TRUE, prob = c(1-p, p))
  mean(x)
})
mean(x_hat_2)
sd(x_hat_2)

```

```{r}
B <- 100
N <- 1000
p <- 0.45
x_hat_3 <- replicate(B, {
  x <- sample(c(0,1), size = N, replace = TRUE, prob = c(1-p, p))
  mean(x)
})
mean(x_hat_3)
sd(x_hat_3)

```

```{r}
B <- 100
N <- 10
p <- 0.45
x_hat_4 <- replicate(B, {
  x <- sample(c(0,1), size = N, replace = TRUE, prob = c(1-p, p))
  mean(x)
})
mean(x_hat_4)
sd(x_hat_4)
``` 



 
 
 
 