---
title: "R Notebook_Ch9_JYChang"
output: html_notebook

---
# Chapter 9 Visualizing data distributions

You may have noticed that numerical data is often summarized with the average value. For example, the quality of a high school is sometimes summarized with one number: the average score on a standardized test. Occasionally, a second number is reported: the standard deviation. For example, you might read a report stating that scores were 680 plus or minus 50 (the standard deviation). The report has summarized an entire vector of scores with with just two numbers. Is this appropriate? Is there any important piece of information that we are missing by only looking at this summary rather than the entire list?

Our first data visualization building block is learning to summarize lists of factors or numeric vectors. More often than not, the best way to share or explore this summary is through data visualization. The most basic statistical summary of a list of objects or numbers is its distribution. Once a vector has been summarized as a distribution, there are several data visualization techniques to effectively relay this information.

In this Chapter, we first discuss properties of a variety of distributions and how to visualize distributions using a motivating example of student heights. We then discuss the ggplot2 geometries for these visualizations in Section 9.16.

# 9.1 Variable types
We will be working with two types of variables: categorical and numeric. Each can be divided into two other groups: categorical can be ordinal or not, whereas numerical variables can be discrete or continuous.

When each entry in a vector comes from one of a small number of groups, we refer to the data as categorical data. Two simple examples are sex (male or female) and regions (Northeast, South, North Central, West). Some categorical data can be ordered even if they are not numbers per se, such as spiciness (mild, medium, hot). In statistics textbooks, ordered categorical data are referred to as ordinal data.

Example of numerical data are population sizes, murder rates, and heights. Some numerical data can be treated as ordered categorical. We can further divide numerical data into continuous and discrete. Continuous variables are those that can take any value, such as heights, if measured with enough precision. For example, a pair of twins may be 68.12 and 68.11 inches respectively. Counts, such as population sizes, are discrete because they have to be round numbers.

Keep in mind that discrete numeric data can be considered ordinal. Although this is technically true, we usually reserve the term ordinal data for variables belonging to a small number of different groups, with each group having many members. In contrast, when we have many groups with few cases in each group, we typically refer to them as discrete numerical variables. So, for example, the number of packs of cigarettes a person smokes a day, rounded to the closest pack, would be considered ordinal, while the actual number of cigarettes would be considered a numerical variable. But, indeed, there are examples that can be considered both numerical and ordinal when it comes to visualizing data.

# 9.7 Exercises
```{r}
library(dplyr)
library(ggplot2)
library(dslabs)
data (murders)
```

1. 
2. B
3. B
4. C
5. D
6. D
7. D
8. A
9. B
10. D

# 9.8  The normal distribution

Histograms and density plots provide excellent summaries of a distribution. But can we summarize even further? We often see the average and standard deviation used as summary statistics: a two-number summary! To understand what these summaries are and why they are so widely used, we need to understand the normal distribution.

The normal distribution, also known as the bell curve and as the Gaussian distribution, is one of the most famous mathematical concepts in history. A reason for this is that approximately normal distributions occur in many situations, including gambling winnings, heights, weights, blood pressure, standardized test scores, and experimental measurement errors. There are explanations for this, but we describe these later. Here we focus on how the normal distribution helps us summarize data.

Rather than using data, the normal distribution is defined with a mathematical formula. For any interval  
(a,b), the proportion of values in that interval can be computed using this formula:

The fact that the distribution is defined by just two parameters implies that if a dataset is approximated by a normal distribution, all the information needed to describe the distribution can be encoded in just two numbers: the average and the standard deviation. We now define these values for an arbitrary list of numbers.

For a list of numbers contained in a vector x, the average is defined as:

```{r}
m <- sum(x) / length(x)
s <- sqrt(sum((x-mu)^2) / length(x))
index <- heights$sex=="Male"
x <- heights$height[index]
m <- mean(x)
s <- sd(x)
c(average = m, sd = s)
#> average      sd 
#>   69.31    3.61
```
# 9.9 Standard units


```{r}
z <- scale(x)
mean(abs(z) < 2)
#> [1] 0.95
```

# 9.10 Quantile-quantile plots

```{r}
pnorm(-1.96)
#> [1] 0.025
qnorm(0.975, mean = 5, sd = 2)
#> [1] 8.92
mean(x <= 69.5)
#> [1] 0.515
p <- seq(0.05, 0.95, 0.05)
sample_quantiles <- quantile(x, p)
theoretical_quantiles <- qnorm(p, mean = mean(x), sd = sd(x))
qplot(theoretical_quantiles, sample_quantiles) + geom_abline()
sample_quantiles <- quantile(z, p)
theoretical_quantiles <- qnorm(p) 
qplot(theoretical_quantiles, sample_quantiles) + geom_abline()
heights %>% filter(sex=="Male") %>%
  ggplot(aes(sample = scale(height))) + 
  geom_qq() +
  geom_abline()
```

# 9.16 ggplot2 geometries
# 9.16.1 Barplots
To generate a barplot we can use the geom_bar geometry. The default is to count the number of each category and draw a bar. Here is the plot for the regions of the US.
```{r}
murders %>% ggplot(aes(region)) + geom_bar()
data(murders)
tab <- murders %>% 
  count(region) %>% 
  mutate(proportion = n/sum(n))
tab
#> # A tibble: 4 x 3
#>   region            n proportion
#>   <fct>         <int>      <dbl>
#> 1 Northeast         9      0.176
#> 2 South            17      0.333
#> 3 North Central    12      0.235
#> 4 West             13      0.255
tab %>% ggplot(aes(region, proportion)) + geom_bar(stat = "identity")
```

# 9.16.2 Histograms
```{r}
heights %>% 
  filter(sex == "Female") %>% 
  ggplot(aes(height)) + 
  geom_histogram()
heights %>% 
  filter(sex == "Female") %>% 
  ggplot(aes(height)) + 
  geom_histogram(binwidth = 1)
heights %>% 
  filter(sex == "Female") %>% 
  ggplot(aes(height)) +
  geom_histogram(binwidth = 1, fill = "blue", col = "black") +
  xlab("Male heights in inches") + 
  ggtitle("Histogram")
```

# 9.16.3 Density plots 
```{r}
heights %>% 
  filter(sex == "Female") %>%
  ggplot(aes(height)) +
  geom_density()
heights %>% 
  filter(sex == "Female") %>%
  ggplot(aes(height)) +
  geom_density(fill="blue")
heights %>% 
  filter(sex == "Female") + 
  geom_density(fill="blue", adjust = 2)
```

# 9.16.4 Boxplots
# 9.16.5 QQ-plots
```{r}
heights %>% filter(sex=="Male") %>%
  ggplot(aes(sample = height)) +
  geom_qq()
params <- heights %>% filter(sex=="Male") %>%
  summarize(mean = mean(height), sd = sd(height))

heights %>% filter(sex=="Male") %>%
  ggplot(aes(sample = height)) +
  geom_qq(dparams = params) +
  geom_abline()
heights %>% 
  filter(sex=="Male") %>%
  ggplot(aes(sample = scale(height))) + 
  geom_qq() +
  geom_abline()
```

# 9.16.6 Images
```{r}
x <- expand.grid(x = 1:12, y = 1:10) %>% 
  mutate(z = 1:120) 
x %>% ggplot(aes(x, y, fill = z)) + 
  geom_raster()
x %>% ggplot(aes(x, y, fill = z)) + 
  geom_raster() + 
  scale_fill_gradientn(colors =  terrain.colors(10))
```

# 9.16.7  Quick plots

```{r}
x <- heights %>% 
  filter(sex=="Male") %>% 
  pull(height)
qplot(x)
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
qplot(sample = scale(x)) + geom_abline()
heights %>% qplot(sex, height, data = .)
heights %>% qplot(sex, height, data = ., geom = "boxplot")
qplot(x, geom = "density")
qplot(x, bins=15, color = I("black"), xlab = "Population")
```