---
title: "R Notebook_Ch4-4.3_JYChang"
output: html_notebook

---
Chapter 4 ~ 4.3
JY Chang

Programming basics

4.1 Conditional expressions

-Flow control; if-else statement

Ex1
```{r}
a <- 0

if(a!=0){
  print(1/a)
} else{
  print("No reciprocal for 0.")
}
#> [1] "No reciprocal for 0."
```

Ex2 
```{r}
library(dslabs)
data(murders)
murder_rate <- murders$total / murders$population*100000
ind <- which.min(murder_rate)

if(murder_rate[ind] < 0.5){
  print(murders$state[ind]) 
} else{
  print("No state has murder rate that low")
}
#> [1] "Vermont"
if(murder_rate[ind] < 0.25){
  print(murders$state[ind]) 
} else{
  print("No state has a murder rate that low.")
}
#> [1] "No state has a murder rate that low."
```

-ifelse statement

ifelse(true,false)

Ex1
```{r}
a <- 0
ifelse(a > 0, 1/a, NA)
#> [1] NA
```

This function works on vectors. It examines each entry of the logical vector and returns elements from the vector provided in the second argument, if the entry is TRUE, or elements from the vector provided in the third argument, if the entry is FALSE.

Ex2
```{r}
a <- c(0,1,2,-4,5)
result <- ifelse(a > 0, 1/a, NA)

```

Here is an example of how this function can be readily used to replace all the missing values in a vector with zeros:
Ex3
```{r}
data(na_example)
no_nas <- ifelse(is.na(na_example), 0, na_example) 
sum(is.na(no_nas))
#> [1] 0
```

Two other useful functions are any and all. The any function takes a vector of logicals and returns TRUE if any of the entries is TRUE. The all function takes a vector of logicals and 
returns TRUE if all of the entries are TRUE.
Ex4
```{r}
z <- c(TRUE, TRUE, FALSE)
any(z)
#> [1] TRUE
all(z)
#> [1] FALSE
```


4.2 Defining functions

Repeat of same operations --> new function

Ex1
```{r}
avg <- function(x){
  s <- sum(x)
  n <- length(x)
  s/n
}

x <- 1:100
identical(mean(x), avg(x))
#> [1] TRUE

s <- 3
avg(1:10)
#> [1] 5.5
s
#> [1] 3
```


- function definition
Ex2
```{r}
my_function <- function(VARIABLE_NAME){
  perform operations on VARIABLE_NAME and calculate VALUE
  VALUE
}
```

-function with multiple arguments
Ex3
```{r}
avg <- function(x, arithmetic = TRUE){
  n <- length(x)
  ifelse(arithmetic, sum(x)/n, prod(x)^(1/n))
}
```


4.3 Namespaces

-load several add-on packages for some of your analysis 
-possibly two packages use the same name for two different functions.
-and often these functions do completely different things.
-These function live in different namespaces.
-R will follow a certain order when searching for a function in these namespaces. 
-You can see the order by typing:
```{r}
search()
```
-So what if we want to use the stats filter instead of the dplyr filter but dplyr appears first in the search list? 
-You can force the use of a specific name space by using double colons (::) like this:
```{r}
stats::filter
```
-If we want to be absolutely sure we use the dplyr filter we can use
```{r}
dplyr::filter
```
-If we want to use a function in a package without loading the entire package, we can use the double colon as well.
